Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-11-16 special-issue-ci/Response to reviewers.tex

We would like to thanks this reviewer for the helpful comments and
suggestion made. We expect to have addressed them properly, but in any
case are grateful because it has allowed us to improve the quality of
our work.

\section{Response to Reviewer \#2}

\begin{quote}
Contribution:
This paper describes a client-server architecture for distributed evolutionary algorithms named NodIO.
The implementation of the architecture is written entirely in JavaScript, in order to support persistent,
asynchronous, distributed evolutionary algorithms running in the browser.
Additionally, the paper presents the results of several experiments performed to validate the proposal and
to measure the performance improvement generated by the massive parallel computing.
I think that the contribution is quite relevant, since it provides a computation platform that is flexible,
extensible and potentially massive. Personally, I consider it one of the fresher and most promising
proposals in the context of metaheuristic optimization frameworks lately. Hence, our global
recommendation of acceptance with minor changes.
\end{quote}

Thanks for your appreciation.

\begin{quote}
However, the paper still has plenty of room for improvement in three main areas: the experimentation, the
organization and description of the platform, and the presentation and writing style.
Regarding the experimentation, I miss:
-  a section discussing the threats to validity of the experiments

\end{quote}

We have inserted a paragraph at the end of the section that describes
the experimental setup describing what are the challenges that this
kind of voluntary framework must overcome. Another paragraph analyzing
the security from the point of view of the framework used has been
also added. A new paragraph on the security challenges has also been
inserted, third paragraph in this section, to clarify our integral and
pragmatic approach to security, which so far has worked perfectly.

\begin{quote}
-  an experiment with a more difficult problem (please use the well-known benchmark problems
described in literature).
\end{quote}

A experiment using the well known Rastrigin function has been
performed and analyzed.

\begin{quote}
-  A comparison in similar conditions of the two implementations of the architecture (the algorithms
used in the experiments are slightly different, but this could lead to dramatic changes in
performance, and consequently, invalidate the comparison performed in the paper). This would
enable a measurement of the actual improvement in performance obtained by the use of web
workers (not that generated due to a different algorithm)
\end{quote}

This has been included as the last subsection of the paper.

\begin{quote}
Regarding the organization of the paper, some important details regarding the improved framework
design named is described in the experimentation section. I would reorganize the contents to describe the
improved in detail in section 3, and focus on the experiments, its
results and the discussion in section 4.
\end{quote}

In fact, reorganizing the paper increased the chance of introducing
new issues in this new version. We have preferred to keep it the way
it is, although I thank you for the suggestion

\begin{quote}
Another issue is that some details are described twice, for instance, the advantages of the use of web
workers. Fixing this issue can reduce the paper size and improve its
readability.
\end{quote}

We have eliminated duplications in this new version. Thanks for
noticing it.

\begin{quote}
Regarding the presentation:
-  Figures should be readable even in gray-scale, so do not use red and blue to denote different
elements, use a different shade or texture.
-  The axis labels and legends in most of the figures are unreadable, please increase its font size.
-  There are some typos and phrases that need rewriting.
\end{quote}

We have rewritten and revised extensively the paper, and tried to find
a way (and failed) to make it more readable for reviewers. For the time being and
during the revision phase, we prefer to keep it that way, but we will
change it to grayscale if it is needed for the final version.

### 2015-11-16 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Adding threats to the system
Closes #35

%---------------------------------------------------------------
\section{Description of the system}
\label{sec:description}

@@

As a volunteer based system, there are some issues worth 
noting \cite{sarmenta2001volunteer,web:BOINC} :\begin{enumerate}
\item Volunteers are anonymous, only the IP address sent in 
each request is known.
\item As anonymous entities, volunteers are not accountable. 
If a volunteer misbehaves in some way the provider cannot 
prosecute or discipline the volunteer. In the case of JavaScript,
this issue is agravated because even if it is obfuscated, sorce code
and data is sent to clients. 
\item Participants must also trust the application provider, 
regarding invation of privacy, the integrity of their devices 
and intended application. 
\end{enumerate}
In this work security aspects related to sabotage and 
trust management (e.g. \cite{domingues2007sabotage}) have not been addressed.

@@

Figure \ref{fig:system} describes the general system architecture and
algorithm behavior. Different web technologies, such as JQuery or {\tt
  Chart.js} have
been used to build the elements of the system.
\begin{figure}[!t]
\centering
\includegraphics[width=4in]{system.pdf}
\caption{Description of the proposed system. Clients execute a JS EA
  in the browser, which, every 100 generations, sends the best
  individual and receives a random one from the server. }
\label{fig:system}
\end{figure}

The researcher only has to change a function, which can be written in
JavaScript or in any other language with cross-compilation to
JavaScript \cite{web:compilersjs} to solve a different
problem. In this case the classical Trap function \cite{Ackley1987} has been
used as well as the Rastrigin floating-point optimization
problem. JavaScript is a functional language and declaring a different 
function and handing at the creation of the algorithm object, called
{\tt Classic} is the only change needed to work with a different
problem. Let us see how this system addresses the different challenges
outlined in the State of the Art (Section \ref{sec:soa}):\begin{itemize}
\item {\em Scalability} is provided via the use of a lightweight and
  high-performance, single-threaded, server based in Node.js and
  Express.js. Although this single server is a bottleneck since it
  will eventually saturate, the fact that it runs as a single thread
  allows the service of many requests. In fact, a limit with the
  number of simultaneous requests will be reached, but so far it has
  not been found, unlike what we found in our previous systems, DCoR,
  \cite{gecco07:workshop:dcor}, which had a low scalability. 
\item The system is {\em heterogeneous} since it does not need any
  performance, operating system or even browser requirement (as long
  as Javascript is enabled): anyone
  visiting the page, even from mobile devices, can load the algorithm.
  In case the browser does not support HTML5 Web workers (e.g., 
  Windows Phone or some version of Android) or their use
  has been disabled, a basic version of NodIO can be used.
\item {\em Fault tolerance} is always an issue, and in this case, the
  single point of failure would be the server: the system, as such,
  would break down if the server fails. However, the individual
  islands in every browser would continue running, and having access
  to just one of them would allow the local algorithm to proceed. In
  fact, the island does not need the server to run: it runs locally if
  needed, with the only exception that it is obviously unable to
  communicate with the rest of the islands.
\item {\em Adaptiveness} is achieved simply through the autonomous
  operation of every individual island with out any synchronization
  mechanism. The islands in the system are, in fact, unaware of each
  other, communicating only through the server. This architecture 
  has been tested in other high demand systems with success, but 
  in the case of NodIO, additional experiments are need to asses 
  the scalability of the system. 
\item Since the algorithm runs in the browser, {\em safety} is
  achieved through its sandbox mechanisms. The user is thus assured
  that there is no unsafe access either to their local files or even
  to more resources that the browser should be allotted.
\item Running the algorithm is just a matter of loading the page,
  which means that operation is totally {\em anonymous}. For the same
  reason, {\em ease of use} is optimal, being as easy as simply
  clicking on an URL and available to anyone with access to a browser.
\item {\em Reasonable performance} is not ensured. In fact, we should
  make sure that there is a reasonable amount of clients over which
  the performance achieved is better than what you would obtain in
  your own desktop system. If this is not the case, it is a pointless
  academic exercise.
\end{itemize}

This last point is what we will try to prove in the next section.

>>>>

%---------------------------------------------------------------
\section{Description of the framework}
\label{sec:description}

@@

As a volunteer based system, there are some issues worth 
noting \cite{sarmenta2001volunteer,web:BOINC} :\begin{enumerate}
\item Volunteers are anonymous, only the IP address sent in 
each request is known.
\item As anonymous entities, volunteers are not accountable. 
If a volunteer misbehaves in some way the provider cannot 
prosecute or discipline the volunteer. In the case of JavaScript,
this issue is agravated because even if it is obfuscated, sorce code
and data is sent to clients. 
\item Participants must also trust the application provider, 
regarding invation of privacy, the integrity of their devices 
and intended application. 
\end{enumerate}

Two kind of threats are possible in this environment; first, since the
API is open and it is an open source system, it is relatively easy to
sabotage the system, as indicated in \cite{domingues2007sabotage}, by
crafting a fake request which, for instance, assigns a fake fitness to
a particular chromosome. Another two different attacks are also
possible: a denial of service attack as well as penetration
testing. We have approached this threat in a pragmatic way: all
sources, client and server, for the application, are released
openly. We also release as open access experiments performed
\cite{DBLP:journals/corr/GuervosG15} and all data is also, after
cleaning it from personal information, released as open data. This
builds trust between the user and the scientist, which indeed has an
impact on the performance of the system, since there is no need to
include cheating checks or other functions that would degrade the
performance of the system. This socio-technical approach is, indeed,
coherent with our whole approach to the study of a volunteer
distributed system which can only be study approaching it from the
technical as well as the social, as in social media, angle.

@@

Figure \ref{fig:system} describes the general system architecture and
algorithm behavior. Different web technologies, such as JQuery or {\tt
  Chart.js} have
been used to build the elements of the system.
\begin{figure}[!t]
\centering
\includegraphics[width=4in]{system.pdf}
\caption{Description of the proposed system. Clients execute a JS EA
  in the browser, which, every 100 generations, sends the best
  individual and receives a random one from the server.}
\label{fig:system}
\end{figure}

The researcher only has to change a function, which can be written in
JavaScript or in any other language with cross-compilation to
JavaScript \cite{web:compilersjs} to solve a different
problem. In this case the classical Trap function \cite{Ackley1987} has been
used as well as the Rastrigin floating-point optimization
problem. JavaScript is a functional language and declaring a different 
function and handing at the creation of the algorithm object, called
{\tt Classic} is the only change needed to work with a different
problem. Let us see how this system addresses the different challenges
outlined in the State of the Art (Section \ref{sec:soa}):
\begin{itemize}
\item {\em Scalability} is provided via the use of a lightweight and
  high-performance, single-threaded, server based in Node.js and
  Express.js. Although this single server is a bottleneck since it
  will eventually saturate, the fact that it runs as a single thread
  allows the service of many requests. In fact, a limit with the
  number of simultaneous requests will be reached, but so far it has
  not been found, unlike what we found in our previous systems, DCoR,
  \cite{gecco07:workshop:dcor}, which had a low scalability. 
\item The system is {\em heterogeneous} since it does not need any
  performance, operating system or even browser requirement (as long
  as Javascript is enabled): anyone
  visiting the page, even from mobile devices, can load the algorithm.
  In case the browser does not support HTML5 Web workers (e.g., 
  Windows Phone or some version of Android) or their use
  has been disabled, a basic version of NodIO can be used.
\item {\em Fault tolerance} is always an issue, and in this case, the
  single point of failure would be the server: the system, as such,
  would break down if the server fails. However, the individual
  islands in every browser would continue running, and having access
  to just one of them would allow the local algorithm to proceed. In
  fact, the island does not need the server to run: it runs locally if
  needed, with the only exception that it is obviously unable to
  communicate with the rest of the islands.
\item {\em Adaptiveness} is achieved simply through the autonomous
  operation of every individual island with out any synchronization
  mechanism. The islands in the system are, in fact, unaware of each
  other, communicating only through the server. This architecture 
  has been tested in other high demand systems with success, but 
  in the case of NodIO, additional experiments are need to asses 
  the scalability of the system. 
\item Since the algorithm runs in the browser, {\em safety} is
  achieved through its sandbox mechanisms. The user is thus assured
  that there is no unsafe access either to their local files or even
  to more resources that the browser should be allotted.
\item Running the algorithm is just a matter of loading the page,
  which means that operation is totally {\em anonymous}. For the same
  reason, {\em ease of use} is optimal, being as easy as simply
  clicking on an URL and available to anyone with access to a browser.
\item {\em Reasonable performance} is not ensured. In fact, we should
  make sure that there is a reasonable amount of clients over which
  the performance achieved is better than what you would obtain in
  your own desktop system. If this is not the case, it is a pointless
  academic exercise.
\end{itemize}

There are many different ways to validate a framework that intends to
address these challenges. Some of them are included by design: it is
anonymous, since no accounts are needed to participate in the
experiment, it is safe for the user, since we are using a the browser
black box, and it is also heterogeneous. We have designed the
experiments so that they show adaptability to different types of
experiments and clients, scalability with the number of clients and
the problem size, and also measured its performance comparing it with
a baseline configuration, with which, as indicated above, is
compatible and can be used in a complementary way. Fault tolerance
will be experimented by dropping the server and check the continuity
of the experiments.

All these points will be established in the next two sections,
starting with the baseline, single worker system next.
